---
title: "p8105_hw2_brm2150"
author: "Brooklynn McNeil"
date: "2024-09-24"
output:
  html_document: default
  pdf_document: default
---

This is an R markdown document for the homework assignment #2 for the p8105 Data Science 1 Class.

Load Packages

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(gt)
```

## Problem 1

## Problem 2

Load Trash Wheel Data from excel and combine Mr. Trash Wheel, Professor Trash Wheel, and Gwynda Trash Wheel.

```{r}
mr.trashwheel_df = 
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N653", na = "") |>
  janitor::clean_names() |>
  mutate(sports_balls = round(sports_balls)) |>
  mutate(sports_balls = as.integer(sports_balls)) |>
  mutate(year = as.double(year)) |>
  mutate(trash_wheel = "mr")

mr.trashwheel_df

prof.trashwheel_df =  
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", sheet = "Professor Trash Wheel", range = "A2:M120", na = "") |>
  janitor::clean_names() |>
  mutate(trash_wheel = "prof") 

prof.trashwheel_df

gwyn.trashwheel_df = 
  read_excel("data/202409 Trash Wheel Collection Data.xlsx", sheet = "Gwynnda Trash Wheel", range = "A2:L265", na = "") |>
  janitor::clean_names() |>
  mutate(trash_wheel = "gwyn") 

gwyn.trashwheel_df

trashwheel_df = 
  bind_rows(mr.trashwheel_df,prof.trashwheel_df, gwyn.trashwheel_df)

trashwheel_df
```

The trash wheel in these 3 locations collected on average `r trashwheel_df |> pull(plastic_bottles) |> mean(na.rm = TRUE) |> round()` plastic bottles per dumpster. The trash wheel data were collected between `r trashwheel_df |> pull(year) |> range()`. Professor Trash Wheel collected a total weight of `r trashwheel_df|>filter(trash_wheel == "prof")|> pull(weight_tons) |> sum()` tons of trash. The total number of cigarette butts that was collected by Gwynda in June of 2022 was `r trashwheel_df |> filter(trash_wheel == "gwyn", month == "June", year == "2022") |> pull(cigarette_butts) |> sum()`.

## Problem 3

Importing and wrangling the data sets from the Great British Bakeoff show. To create the combined data set I took the following steps:

1.  `read_csv` to import the data.
    1.  some data sets needed `nas` to be specified.
    2.  some data sets needed rows skipped at the beginning of the file.
2.  `clean_names` to transform all of the variable names to snake case.
3.  In `bakers_df` the `baker_name` variable needed to have the last name removed so it could be used to join with the , so I used `separate` and `select` to remove the last name.
4.  I selected only the unique columns in `bakes_df` and `results_df` so that it didn't cause issues when joining the data sets together.
5.  I then created a new data frame `gbb_df` that joined together the 3 data sets using the `full_join` function.I also made sure to join by `baker`, `series` and `episode` to make sure the data was aligned and no duplicates were made.
6.  I also found that a few data was missing from the final data frame because it had missing values for all of the variables. This was checked using the `anti-join` function.
7.  I used `tidyverse` and the `pipe` function to achieve this data wrangling.


It seems that "Jo" is being left out and is maybe named "Joanne" in other tables. I am choosing to not include them in the final table because it is unclear if they are the same person or not. 

```{r}
bakers_df =
  read_csv("data/bakers.csv") |>
  janitor::clean_names() |>
  separate(baker_name, into = c("baker", "last_name"), sep = " ")

bakers_df

bakes_df = 
  read_csv("data/bakes.csv", na = "N/A") |>
  janitor::clean_names()

bakes_df

results_df = 
  read_csv("data/results.csv", skip = 2, na = "NA") |>
  janitor::clean_names()

results_df

gbb_df = 
  results_df |>
  full_join(bakes_df, by = c("baker", "series","episode")) |>
  full_join(bakers_df, by = c("baker", "series")) 

gbb_df

anti_join(bakes_df, bakers_df, by = "baker")
anti_join(bakes_df, results_df, by = "baker")
anti_join(results_df, bakers_df, by = "baker")
```

Creating a table showing the winner of each episode in Season 5 through 10

```{r}
gbb_df |>
  filter(result == "STAR BAKER"|result == "WINNER", series >=5) |>
  select(c(baker, series, episode, result)) |>
  arrange(series, episode) |>
  gt() |>
  tab_header(
    title = "Great British Bakeoff Episode Winners",
    subtitle = "Episode Winners from Season 5 through 10"
  )
```
In season 5 it looks like there was an upset because Richard won the last 3 episodes before the finale but Nancy took the cake *literally*. Nadiya powered through the last 3 episodes of season 6 to become the winner. Rahul looks like he started strong and then lost some steam, but came back to win the season in the end. Season 10 was probably the most surprising because the winner, David, was never crowned star baker in an episode before the finale. 


Import, tidy, and wrangle the *viewers* data.
```{r}
viewers_df = 
  read_csv("data/viewers.csv", na = "NA") |>
  janitor::clean_names() |>
  pivot_longer(
    cols = series_1:series_10, 
    names_to = "series", 
    values_to = "viewership",
    names_prefix = "series_") |>
  drop_na()

head(viewers_df, n=10)
```
The average viewership for season 1 was `r viewers_df |> filter(series == "1")|> pull(viewership) |> mean()` and for season 5 was `r viewers_df |> filter(series == "5")|> pull(viewership) |> mean()`
